---


title: "Project1_course5"
author: "aomelanuk"
date: '2018-08-30'
output:
  html_document: default
  pdf_document: default
  word_document: default
---


Github repo for the Course: [Reproducible Research](https://github.com/aomelanuk/course_5_project1) 

---


### 1.    Synopsis

The goal of this work collect a large amount of data about personal movement using activity monitoring and handle it for analisys

---


### 2. Checking environment

loading needable librariesfor data processing in this work I will use next libraries: dplyr, ggplot2

```{r setup, include=TRUE}
      

library("ggplot2")
library("dplyr")
library("lubridate")
```
---


### 3.1 Loading data
Checking exist data on the PC, if not exist dovnload it from source (https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip)

```{r }
if(!(file.exists("activity.csv")))
{
           if(!(file.exists("repdata_data_activity.zip")))  
           {
                      url<-"https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
                      download.file(url,"repdata_data_activity.zip")
           }
           unzip("repdata_data_activity.zip",overwrite = TRUE)
}
my_data<-tbl_df(read.csv("activity.csv",na.strings = "NA",colClasses = "character"))
```
for loading I use function read.csv with parameters 
for data handling I use variable my_data
As result I have data for next analisys, but all variables are chracter and next step I transform variable to useful types

```{r}
my_data$date<-ymd(my_data$date)
my_data$interval<-as.integer(my_data$interval)
my_data$steps<-as.integer(my_data$steps)

```
---


### 3.2 Histogram of the total number of steps taken each day

For results mean, average I create new data frame with this results by day
I plot histogram using new table my_data_h with mean, median, and sum

```{r}
my_data_h<-my_data%>%select(date,steps)%>%group_by(date)%>%summarise_all(c(funs(mean),funs(sum)))
head(my_data_h)
hist(my_data_h$sum,breaks = 26,col = "red", main = "Histogram number of steps", xlab = "steps quantity")

```
---


###3.3 Mean and median number of steps taken each day
for calculate this data I use data frame my_data_h

```{r}
median_steps=median(my_data_h$sum,na.rm = TRUE)
print(paste("median steps quantity", median_steps))
mean_steps=mean(my_data_h$sum,na.rm = TRUE)
print(paste("Mean steps quantity", mean_steps))

```
---


###3.4 Time series plot of the average number of steps taken

```{r}
mean_interval<-my_data%>%
      filter(!is.na(steps))%>%
      select(interval,steps)%>%
      group_by(interval)%>%
      summarise_all(funs(mean)) 
plot(x=mean_interval$interval,y=mean_interval$steps,
     type = "l",main = "Time series plot of the average number of steps taken",
     xlab = "interval", ylab = "average steps", col="red")                          

```
---


### 3.5 The 5-minute interval that, on average, contains the maximum number of steps

```{r}
top_int<-my_data%>%filter(!is.na(steps))%>%select(interval,steps)%>%
      group_by(interval)%>%summarise_all(funs(mean))%>%
      arrange(desc(steps))%>%head(1)
print(top_int)

```


---

###3.6 Code to describe and show a strategy for imputing missing data
in first step I lookin rows with NA data
```{r}

print(paste("Quantity NA rows",my_data%>%filter(is.na(steps))%>%nrow))

```
days when steps data NA

```{r}

na_date<-my_data%>%filter(is.na(steps))%>%select(date)%>%unique()
print("date with NA data")
print(as.character(na_date$date))

```

for inserting data in rows with NA I use data frame mean_interval (was created before)
```{r}
my_data[is.na(my_data$steps),"steps"]<-mean_interval$steps

```

---

### 3.7 Histogram of the total number of steps taken each day after missing values are imputed

```{r}
my_data_h<-my_data%>%select(date,steps)%>%
      group_by(date)%>%summarise_all(c(funs(mean),funs(sum)))
head(my_data_h)
hist(my_data_h$sum,breaks = 26,col = "red", 
     main = "Histogram number of steps,without NA", xlab = "steps quantity")
```
---

### 3.8 Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
my_data<-my_data%>%mutate(weekday = wday(date))


```{r, fig.height = 9}

my_data<-my_data%>%mutate(weekday = wday(date))
mean_interval_w<-my_data%>%
      filter(weekday==c(1,7))%>%
      select(interval,steps)%>%
      group_by(interval)%>%
      summarise_all(funs(mean))
mean_interval<-my_data%>%
      filter(weekday==c(2:6))%>%
      select(interval,steps)%>%
      group_by(interval)%>%
      summarise_all(funs(mean))
par(mfrow=c(2,1))   

plot(x=mean_interval$interval,y=mean_interval$steps,
     type = "l",main = "Time series plot of the average number of steps taken on weekdays", xlab = "interval", ylab = "average steps", col="green")    
plot(x=mean_interval$interval,y=mean_interval$steps,
     type = "l",
     main = "Time series plot of the average number of steps taken on other days",
     xlab = "interval", ylab = "average steps", col="blue") 
```
